%TCIDATA{Version=4.00.0.2321}
%TCIDATA{LaTeXparent=0,0,sw-edit.tex}
                      
%TCIDATA{ChildDefaults=chapter:1,page:1}
%
% Time-stamp: <05/03/09 22:01:22 vilhuber>

%\section{Forming Aggregated Estimates}
%\label{sec:aggregate}


\subsection{What are the QWI statistics?}
\label{sec:qwi:what}

[ describe variables / statistics here ]


\subsection{Computing the statistics}
\label{sec:qwi:compute}


Aggregating the QWI data is a four step process, which can be summarized as
follows:

\begin{itemize}
\item[\ ] 

\begin{enumerate}
\item The basic variables, as discussed above, are created for each
employment history (PIK\index{PIK}-SEIN\index{SEIN}-SEINUNIT\index{SEINUNIT} pair) and for every quarter
that the pair exits. Note that each job history stems from an implicate of
the U2W, and
is weighted accordingly. 


\item The QWI system sums for each employer's workplace the following variables:
beginning-of-period employment, end-of-period employment, accessions, new
hires, recalls, separations, full-quarter employment, full-quarter
accessions, full-quarter new hires, total earnings of full-quarter
employees, total earnings of full-quarter accessions, and total earnings of
full-quarter new hires. Job creations, job destructions, and net job flows
are estimated for each workplace using the beginning and end of quarter
employment estimates for that workplace. The first-layer of
disclosure-proofing is also applied at this step.

\item The workplace-level variables in the list above are summed over the
relevant aggregating unit (county or SIC) for each quarter. Average
earnings of full-quarter employees, full-quarter accessions, and
full-quarter new hires are estimated by taking the ratio of total earnings
of the relevant category to the total number of individuals in that
category. \ For example, avearge earnings of full-quarter men ages 55-64 for
a given year, quarter and county is the ratio of total earnings of
full-quarter men ages 55-64 to the number of full-quarter men ages 55-64 in
that year, quarter, and county.

\item The beginning-of-quarter employment for each county or SIC division is
controlled (raked) to the BLS estimate of total county employment in month
one of that quarter from the Covered Employment and Wages series. At this
point the other estimates and the demographic groups are also raked to
preserve the underlying relations among the variables.
\end{enumerate}
\end{itemize}

%\chapter{Worker and Job Flow Analysis\label{cha:jobflow}}
%
%\input{\mypath/jobflow_analysis.tex}
%\input{\mypath/jobflow_analysis_data_consistency.tex}


\section{The Final product}
\label{sec:final}

\subsection{Summary Variable Definitions}

\subsubsection{Timing and Category Variables}

\index{Timing Variables|see{Variables, timing}} 
\index{Category Variables|see{Variables, category}} \mindex{Variables!timing}
\mindex{Variables!category}

Timing and categorical variables are used to describe the population and
time period that the content variables cover. The first such variable is 
\textsf{STATE}\mindex{STATE}, which is the two-digit FIPS%
\index{FIPS} code for the state upon which the employment dynamics estimates
are based. The next two variables (\textsf{YEAR}\mindex{YEAR} and \textsf{%
QUARTER}\mindex{QUARTER}) refer to the calendar year and quarter covered by
the content variables. The \textsf{COUNTY}\mindex{COUNTY} variable
(county-level data file) is the three-digit FIPS%
\index{FIPS} code for the county (within the state). The \textsf{%
SIC\_DIVISION}\mindex{SIC\_DIVISION} variable (sic-division-level data file)
is the one-character SIC\ (1987) major industry group. The \textsf{SEX}%
\mindex{SEX} variable indicates whether the data cover men or women. The 
\textsf{AGEGROUP}\mindex{AGEGROUP} variable indicates which of the eight age
categories the data cover.

\subsubsection{Content Variables}

\index{Content Variables|see{Variables, content}} \mindex{Variables!content}

The quarterly employment estimates for beginning of quarter employment are
contained the variable \textsf{B} and the estimates for end of quarter
employment are found in the variable \textsf{E}. Accessions are reported in
the variable \textsf{A}. New hires are in \textsf{H} and recalls are
reported in \textsf{R}. Separations are reported in the variable \textsf{S}.

Because of the confidentiality protection system used for the Employment
Dynamics Estimates, the estimate of beginning-of-quarter employment for both
sexes (\textsf{SEX}=0) and all age groups (\textsf{AGEGROUP}=0) is exactly
equal to the BLS-published Covered Employment and Wages estimate of
employment on the 12th day of the first month of the quarter for the
relevant geographic and industrial category. For example, in California the
QWI estimate for beginning-of-quarter employment in the entire state in
1999:3 is 14,440,000 (\textsf{B}=14,440,000 for \textsf{STATE}=``06'', 
\textsf{YEAR}=1999, \textsf{QUARTER}=3, \textsf{COUNTY}=``000'' (or \textsf{%
SIC\_DIVISION}=(blank)),\textsf{SEX}=0, \textsf{AGEGROUP}=0), which exactly
equals the BLS CEW estimate for month 1 in 1999:3 for the entire state,
combining all establishment sizes and all ownership categories. \ Similarly,
the QWI estimate of end-of-quarter employment is controlled for the category
both sexes (\textsf{SEX}=0) and all age groups (\textsf{AGEGROUP}=0) to
equal the BLS-published CEW estimate of employment on the 12th day of the
first month of the succeeding quarter. Again, considering California, the
QWI estimate for end-of-quarter employment in 1999:3 is 14,660,000, which
exactly equals the BLS\ CEW estimate for month 1 in 1999:4 for the entire
state (\textsf{E}=14,660,000 for \textsf{STATE}=``06'', \textsf{YEAR}=1999, 
\textsf{QUARTER}=3, \textsf{COUNTY}=``000'' (or \textsf{SIC\_DIVISION}%
=(blank)), \textsf{SEX}=0, \textsf{AGEGROUP}=0). 
%See the tables in Appendix %
%\ref{app:raking_data} for a complete list of the BLS\ series used in this
%control.

Quarterly employment estimates are also provided on a full-quarter basis.
These estimates are reported in the variable \textsf{F}. Full-quarter
accessions are reported in \textsf{FA}. Full-quarter separations are
reported in \textsf{FS}. Full-quarter new hires are in \textsf{H3. }The
raking step of the QWI confidentiality protection system used to disclosure
proof the variables \textsf{B} and \textsf{E} (and related variables) does
not affect the estimates of full-quarter employment and related flows.

Job creations and destructions are reported in the variables \textsf{JC} and 
\textsf{JD}, respectively. Net job flows are reported in the variable 
\textsf{JF}. Full-quarter job creations and destructions are reported in 
\textsf{FJC} and \textsf{FJD}, respectively. Full-quarter net job flows are
in \textsf{FJF}.

Average earnings of full-quarter employees can be found in \textsf{Z\_W3}.
Average earnings of full-quarter new hires are reported in \textsf{Z\_WH3}.

%See the table of contents at the end of this primer for a list of other
%variables and definitions in the QWI data files.

\subsubsection{Status Flag Variables}

\index{Status Flag Variables|see{Variables, status flag}} %
\mindex{Variables!status flag}

Every variable in QWI data files has an associated status flag. These
variables are called \textsf{[varname]\_status}. The status flag variables
are also shown in the contents tables at the end of this primer. The status
flag has three distinct values:

\begin{itemize}
\item[$*$] indicates significant distortion is necessary to preserve
confidentiality

\item[$d$] indicates an estimate is based on $<3$ employees in the at-risk group.

\item[$n$] indicates an estimate is not defined because no employees are in
the relevant category
\end{itemize}

%\newpage

% \subsection{Data structure\label{primer_tables}}
% 
% Appendix~\Vref{app:data_structure} describes the contents of a typical
% output statistics file, in this case for the state of Texas, and aggregated
% to the county level. A similar file exists at the industry aggregation
% level, and the same pair of files is constructed for every available state.
% 
%\input{\mypath/ca_county_v23_fuzzed.tex} \newpage \input{\mypath/ca_county_v23_fuzzed.freq.tex}
%\newpage \input{\mypath/ca_sic_division_v23_fuzzed.tex} \newpage \input{%
%\mypath/ca_sic_division_v23_fuzzed.freq.tex}

% \newpage
% 
% \subsection{Florida}
% 
% \input{\mypath/fl_county_v23_fuzzed.tex} \newpage \input{\mypath/fl_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/fl_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/fl_sic_division_v23_fuzzed.freq.tex}
% 
% \newpage
% 
% \subsection{Illinois}
% 
% \input{\mypath/il_county_v23_fuzzed.tex} \newpage \input{\mypath/il_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/il_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/il_sic_division_v23_fuzzed.freq.tex}
% 
% \newpage
% 
% \subsection{Maryland}
% 
% \input{\mypath/md_county_v23_fuzzed.tex} \newpage \input{\mypath/md_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/md_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/md_sic_division_v23_fuzzed.freq.tex}
% 
% \newpage
% 
% \subsection{Minnesota}
% 
% \input{\mypath/mn_county_v23_fuzzed.tex} \newpage \input{\mypath/mn_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/mn_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/mn_sic_division_v23_fuzzed.freq.tex}
% 
% \newpage
% 
% \subsection{North Carolina}
% 
% \input{\mypath/nc_county_v23_fuzzed.tex} \newpage \input{\mypath/nc_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/nc_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/nc_sic_division_v23_fuzzed.freq.tex}
% 
% \subsection{Texas}
% 
% \input{\mypath/tx_county_v23_fuzzed.tex} \newpage \input{\mypath/tx_county_v23_fuzzed.freq.tex}
% \newpage \input{\mypath/tx_sic_division_v23_fuzzed.tex} \newpage \input{%
% \mypath/tx_sic_division_v23_fuzzed.freq.tex}


%\subsection{Example statistics}
%\label{sec:example_statistics}
%
%Table~\Vref{tab:table_example} shows the extreme case of statistics for a
%small county and for one gender only.%
%%
%%\footnote{The exact county and gender remain anonymous for data
%%  confidentiality reasons.} 
%%
%Two types of confidentiality measures are
%implemented within the same table. Ten cells have been distorted through
%the injection of noise described earlier. Furthermore, one cell has been
%suppressed because the statistics in that cell is based on too few
%individuals.
%
%
%  \input{table_example_statistics}
%  
%  Figure~\Vref{fig:figure_example} presents the data in a different way.
%  The graph shows net job creation rates by county, for both genders, and
%  for the youngest workers. It shows large dispersion across the state, and
%  highlights the value added from joining demographic and firm-level data
%  to form new statistics.
%
%  \input{figure_example_statistics}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "qwi-overview"
%%% End: 
