%TCIDATA{Version=4.00.0.2321}
%TCIDATA{LaTeXparent=0,0,sw-edit.tex}
                      
%
% Time-stamp: <05/04/05 14:03:46 vilhuber>

%\section{Forming Aggregated Estimates}
%\label{sec:aggregate}

The infrastructure files in principle have all the information necessary to
compute the QWI statistics. However, to be able to compute statistics at a
low level of geographical aggregation, we need to be able to place
individuals at specific work locations. And since the focus is on dynamics
-- flows of workers in and out of firms and workplaces -- the economic
definition of a workplace and a firm needs to be well-defined.


\section{Connecting firms intertemporally: the Successor-Predessor File
(SPF)}
\label{sec:spf}

The firm identifier used in all of LEHD's files is a state-specific
account number in that state's unemployment insurance system, used in
particular to collect payroll taxes. These account numbers, here called 'SEIN', can and do
change for a number of reasons, including a simple change in legal form or
a merger. Typically, the separation of a worker from a firm is
identified by a change in the SEIN  on that worker's wage
records. If a firm changes SEINs, but makes no other changes, the
worker would seem to have left the original firm, when in fact his
employment status remains unchanged. Thus, a simple change in account
numbers would lead to the observation of a firm closing, when in fact, all
workers remain employed.

To identify such events, the Successor Predecessor File (SPF) tracks large worker movements between SEINs.
\citet{tp-2003-09} used the SPF for an early analysis in one particular state of
the impact of such an exercise. The SPF provides for a variety of link
characteristics, based on the number of workers leaving an SEIN, in both
absolute and relative terms, and the number of workers entering an SEIN,
again in absolute and relative terms. 

For the QWI, only the strongest links are used to filter out spurious firm
identifier changes: If 80\% of an SEIN's workers (the predecessor) are
observed to move to a single successor, and that successor absorbs 80\% of
its employees from a single predecessor, then all flows between those two
account numbers are filtered out, and treated as if they had never existed.

An evaluation of the impact of the SPF on the aggregate QWI statistics is
currently under way.

Of importance to the Unit-to-Worker impute described in the next section is
a similar measure, computed within an SEIN. For most states and firms within
states, the breakout of units into SEINUNITs is  at the discretion
of the firm, and the firm may decide to change such a breakout. Again, by
following groups of workers as they move between SEINUNITs, spurious
intra-SEIN flows can be detected. 


\section{Allocating workers to workplaces: Unit-to-worker impute (U2W)}
\label{sec:u2w}


Early versions of the QWI (then
called the Employment Dynamics Estimates, EDE), were computed only at the SEIN
level, with employment allocated to a single location per SEIN. This
approach was driven by the absence of workplace information on almost all
state-provided wage records. Only the state of Minnesota requires the
identification of a worker's workplace (SEINUNIT) on a wage records. 

%For the QWI, a more sophisticated approach was needed. In a nutshell, using
%the Minnesota data, a model relating proximity between residence and
%potential workplaces for workers was developed. The parameters from that
%model, which are regularly updated, are then used to impute a worker's
%workplace. This imputation is performed ten times, and all ten implicates
%are later used in the computation of the QWI. 

%\input{impute_description.tex}
\input{impute_description_bryce_edits.tex}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "qwi-overview"
%%% End: 
