%TCIDATA{LaTeXparent=0,0,sw-edit.tex}
                      

% -*- latex -*- 
%
% Time-stamp: <02/03/14 19:06:30 vilhuber> 
%              Automatically adjusted if using Xemacs
%              Please adjust manually if using other editors
%
% glossary.tex
% Responsible: Lars/John 
% Part of QWI_methods.tex
%
% \mindex{} inserts a main entry into the Index
% \index{} inserts a generic entry into the Index
% \Mindex{} inserts a main entry into the Index and prints the word, too
% \aindex{} inserts a generic entry into the Index and prints the word, too

\begin{description}
\item[ACS:] \mindex{ACS} American Community Survey
\item[ARRS:] \mindex{ARRS} Administrative Records Research Staff, part of
  the U.S. Census Bureau's PRED
\item[ASCII:] \mindex{ASCII} American Standard Code for Information
  Interchange, typically used to denote raw text files in PC or Unix environments
\item[BLS:] \mindex{BLS} \href{http://www.bls.gov}{Bureau of Labor Statistics}
\item[CPR:] \mindex{CPR} Census place of residence (file)
\item[CPS:] \mindex{CPS} \href{http://www.bls.gov/cps/home.htm}{Current Population Survey}
\item[CEW:] \mindex{CEW} \href{http://www.bls.gov/cew/home.htm}{Covered Employment and Wages}. Employment statistics program run by BLS in
  conjunction with all states, also known as ES-202. Generally, when used
  in this document, refers to public-use tabulations from the CEW, as
  opposed to the confidential microdata received directly from the states.
\item[DSBU02:] \mindex{DSBU02} A Compaq GS-160 8-processor machine, running
  OSF with  24~GB of RAM, used for processing all QWI data.
\item[ECF:] \mindex{ECF} Employer Characteristics  File 
\item[QWI:] \mindex{QWI} Employment dynamics estimates
\item[EHF:] \mindex{EHF} Employment History File
\item[ES-202:] \mindex{ES-202} The employer-level data files received from the states,
  which originate from the same source files that feed the CEW, are
  typically referred to as ES-202 data. These are the confidential microdata.
\item[ICF:] \mindex{ICF} Individual Characteristics File
\item[IRS:] \mindex{IRS} \href{http://www.irs.gov}{Internal Revenue Service}
\item[LEHD:] \mindex{LEHD} Longitudinal Employer-Household Dynamics, a   program of the U.S. Census Bureau
\item[MOU:] \mindex{MOU} Memorandum of Understanding, also called Data Use
  Agreement
\item[Numident, Census:] \mindex{Numident!Census} an older version of the Personal Characteristics File
\item[Numident, SSA:] \mindex{Numident} Source of some of the information
  on the PCF
\item[OSF:] \mindex{OSF} Variant of UNIX running on Compaq machines
\item[PCF:] \mindex{PCF} Personal Characteristics File, also called Numident
\item[PIK:] \mindex{PIK} Protected Identity Key
\item[PIKizing:] \mindex{PIKizing} Process of assigning PIK
\item[PRED:] \mindex{PRED} Planning, Research and Evaluation Division (home
  of ARRS)
\item[RUN:] \mindex{RUN} Reporting unit number, identifies sub-SEIN units,
  typically an establishment, in the CEW system
\item[SAS:] A statistical software package, by \href{http://www.sas.com}{SAS Inc.}
\item[SEIN:] \mindex{SEIN} State employer identification number, an LEHD
  acronym for account number identifying individual employers in the state
  UI system
\item[SIC:] \mindex{SIC} Standard Industry Classification
\item[SIPP:] \mindex{SIPP} \href{http://www.sipp.census.gov/sipp}{Survey of Income and Program Participation} 
\item[SSA:] \mindex{SSA} \href{http://www.ssa.gov}{Social Security Administration}
\item[SSEL:] \mindex{SSEL} former name of Business Register
\item[SSN:] \mindex{SSN} Social Security Number, assigned by SSA
\item[UI:] \mindex{UI} Unemployment Insurance
\item[UNIX:] \mindex{UNIX} A multi-user, multitasking operating system
  developed at Bell Labs in the early 1970s. Variants include (Compaq's)
  OSF, run on LEHD's computers, Linux, and Solaris. All programming at LEHD
  is done on UNIX systems.
\item[WIA:] \mindex{WIA} Workforce Investment Act
\end{description}


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "QWI_methods"
%%% End: 
