%TCIDATA{LaTeXparent=0,0,sw-edit.tex}
                      
%
% Time-stamp: <05/04/06 13:48:43 vilhuber>


%\section{Post-processing of UI and ES202 files}
%\label{sec:post-processing}

Once received, the UI and ES202 files are standardized.%
%
\footnote{%
The
ES202 files in particular have been received in a bewildering array of physical file
layouts and formats, reflecting the wide diversity in computer systems
installed in state agencies.}%
%
The UI files have been edited for longitudinal
consistency, and the SSN replaced by the Protected Identification Key
(PIK). Beyond that, no further processing has occurred.

The core Infrastructure Files are built from the core input files, and
augmented from a large number of additional Census-internal demographic and
economic (firm) surveys and censuses. 
%
The Employment History File (EHF\index{EHF}) provides a full time-series of
 earnings at all within-state jobs for all time periods covered by the
LEHD data, and activity calendars at a job, SEINUNIT, and SEIN level. 
%
The Individual Characteristics File
(ICF\index{ICF}) provides time-invariant personal characteristics and
some address information.%.
%
\footnote{A time-varying variant of the ICF is under development.}%
%
The Employer Characteristics File (ECF\index{ECF}) provides a complete 
database of firm and establishment characteristics, most of which are
time-varying. 
%
It includes a subset of the data available on the Geocoded Address List
(GAL\index{GAL}), which contains geocoded at the block-level  and
latitude/longitude coordinates for addresses from a large set of
administrative and survey data.
%
We will describe each in detail.


\section{Employment History File\label{cha:ehf}: EHF}
\input{\mypath/ehf_description.tex}

%
%

\label{sec:files}
\section{Individual Characteristics File\label{cha:icf}: ICF}
\input{\mypath/icf_description.tex}

%
%
\section{The Employer Characteristics File\label{cha:ecf}: ECF}
\input{\mypath/ecf_description.tex}
\input{\mypath/ecf_naics.tex}

%
%

\section{The Geocoded Address List\label{cha:gal}: GAL}
\input{\mypath/gal_intro.tex}

%
%



%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "qwi-overview"
%%% End: 
