% $Id: modifications.tex 461 2008-09-03 19:46:44Z vilhu001 $
% $URL: https://repository.vrdc.cornell.edu/private/lv39_Papers/LEHD_definitions_revisions/text/modifications.tex $
% Time-stamp: <08/09/03 15:33:37 vilhuber>


\subsection{Stable jobs}

We currently have the FQ definitions (see equations \ref{eq:fijt},
\ref{eq:a3ijt}, and \ref{eq:s3ijt}), and they are applied to flows as
well. The full-quarter definitions are internally consistent (see
Definitions in Section \ref{sec:identities}), and are
treated much like the point-in-time measures; compare

\begin{definition}
Evolution of end of period employment 

(using $E_{jt - 1}$ instead of
$B_{jt}$ for clarity)
\begin{equation*}
E_{jt} = E_{jt - 1} + A_{jt} - S_{jt}
\end{equation*}
\end{definition}

and

\begin{definition}
Evolution of full-quarter employment 
\begin{equation*}
F_{jt} = F_{jt - 1} + FA{\ }_{jt} - FS_{jt}
\end{equation*}
\end{definition}

and associated flow and creation/destruction identities. Note however that
if we write the evolution of {\it beginning of period employment}, there are
timing differences in the notation:

\begin{definition}
Evolution of beginning of period employment 

\begin{equation*}
B_{jt} = B_{jt - 1} + A_{jt-1} - S_{jt-1}
\end{equation*}
\end{definition}

Thus, for contemporaneous beginning of period employment, it is the flows
from the previous quarter that must be taken into account.

To provide a modified concept of ``stable jobs'' that relate more
``naturally'' to ``non-stable'' flow concepts (separations, accessions), we
will want them to be in sync with equations \ref{eq:s2ijt} and
\ref{eq:a2ijt}:

\paragraph{Accessions to consecutive quarter status}

\mindex{Accessions!to consecutive quarter status}

$\left( a_{2}\right) $: for $qfirst<t<qlast$, individual $i$ transited from
accession to consecutive-quarter status at $j$ at the end of $t$ and the
beginning of $t+1$ (accession in $t$ and still employed at the end of the
quarter)%
\begin{equation*}
a_{2ijt}=\left\{ 
\begin{array}{l}
1,\text{ if }a_{1ijt}=1\;\&\;m_{ijt+1}=1 \\ 
0,\text{ otherwise.}%
\end{array}%
\right.  (\ref{eq:a2ijt})
\end{equation*}

\paragraph{Separations from consecutive quarter status}


$\left( s_{2}\right) $: for $qfirst<t<qlast$, individual $i$ separated from $%
j$ during $t$ with consecutive-quarter status at the start of $t$%
\begin{equation*}
s_{2ijt}=\left\{ 
\begin{array}{l}
1,\text{ if }s_{1ijt}=1\;\&\;m_{ijt-1}=1 \\ 
0,\text{ otherwise.}%
\end{array}%
\right.   (\ref{eq:s2ijt})  
\end{equation*}

Note that equivalently, $a_{2ijt}=1$ if $b_{ijt}=0$ and $e_{ijt}=1$, and
$s_{2ijt}=1$ if $b_{ijt}=1$ and $e_{ijt}=0$. 

Let's start with the above, expressed in terms of the relevant $f$
variable:

\paragraph{Accessions to stable job}


$\left( a_{4}\right) $: for $qfirst<t<qlast-1$, individual $i$ acceded to a
``stable job'' with $j$ during $t$  (accession in $t$ and still employed at
the end of the next quarter)%
\begin{equation}
a_{4ijt}=\left\{ 
\begin{array}{l}
1,\text{ if }a_{1ijt}=1\;\&\;f_{ijt+1}=1 \\ 
0,\text{ otherwise.}%
\end{array}%
\right.  
\end{equation}

or equivalently, $a_{4ijt}=1$ if $f_{ijt+1}=1$ and $b_{ijt}=0$. Note that
by construction, $A_4<A_1$ (and also $A_4<A_2$)

\paragraph{Separations from stable jobs}


$\left( s_{4}\right) $: for $qfirst+1<t<qlast$, individual $i$ separated
from a stable job with $j$ during $t$ 

\begin{equation*}
s_{4ijt}=\left\{ 
\begin{array}{l}
1,\text{ if }s_{1ijt}=1\;\&\;f_{ijt-1}=1 \\ 
0,\text{ otherwise.}%
\end{array}%
\right.   
\end{equation*}


or equivalently, $s_{4ijt}=1$ if $f_{ijt-1}=1$ and $e_{ijt}=0$. Note that
by construction again, $S_4<S_1$ (and also $S_4<S_2$)

\paragraph{Stable jobs}

Stable jobs are jobs that ever achieve FQ status. It is sufficient to check
adjacent quarters, in addition to the contemporaneous one. Equivalently,
stable jobs are jobs that last at least three quarters. It is desirable to
have them be a subset of end-of-quarter jobs to get consistent flows.



\begin{equation}
sj_{ijt}=\left\{ 
\begin{array}{l}
1,\text{ if }e_{ijt}=1 \text{ and }
  \left ( 
      f_{ijt-1} = 1 
      \text{ or } f_{ijt-1} = 1
      \text{ or } f_{ijt+1} = 1
   \right )
   \\ 
0,\text{ otherwise.}%
\end{array}%
\right.   \label{eq:sjijt}
\end{equation}

Essentially, the relationship between stable jobs and full-quarter jobs is
graphically obvious:

\paragraph{Stable jobs active}

\begin{tabular}{cccccc}
  m & 0 & 1 & 1 & 1 & 0 \\
  f & 0 & 0 & 1 & 0 & 0 \\
  e & 0 & 1 & 1 & 0 & 0 \\
 sj & 0 & 1 & 1 & 0 & 0 \\
\hline
 a1 & 0 & 1 & 0 & 0 & 0 \\
 a2 & 0 & 1 & 0 & 0 & 0 \\
 a3 & 0 & 0 & 1 & 0 & 0 \\
 a4 & 0 & 1 & 0 & 0 & 0 \\
\hline
 s1 & 0 & 0 & 0 & 1 & 0 \\
 s2 & 0 & 0 & 0 & 1 & 0 \\
 s3 & 0 & 0 & 0 & 1 & 0 \\
 s4 & 0 & 0 & 0 & 1 & 0 \\
\hline
\end{tabular}
\paragraph{Stable jobs inactive}

\begin{tabular}{cccccc}
  m & 0 & 1 & 1 & 0 & 0 \\
  f & 0 & 0 & 0 & 0 & 0 \\
  e & 0 & 1 & 0 & 0 & 0 \\
 sj & 0 & 0 & 0 & 0 & 0 \\
\hline
 a1 & 0 & 1 & 0 & 0 & 0 \\
 a2 & 0 & 1 & 0 & 0 & 0 \\
 a3 & 0 & 0 & 0 & 0 & 0 \\
 a4 & 0 & 0 & 0 & 0 & 0 \\
\hline
 s1 & 0 & 0 & 0 & 1 & 0 \\
 s2 & 0 & 0 & 0 & 1 & 0 \\
 s3 & 0 & 0 & 0 & 0 & 0 \\
 s4 & 0 & 0 & 0 & 0 & 0 \\
\hline
\end{tabular}

\paragraph{Stable job flow identities}

Note that 

\begin{equation}
SJ_{jt}=sj_{\cdot jt}
\end{equation}

Because it is defined as a subset of $e$, the flows and identities carry
through, with the appropriate modifications:


\paragraph{Net stable job flows}

(change in employment) for establishment $j$ during period $t$%
\begin{equation}
JF_{4jt}=SJ_{jt}-SJ_{jt-1}  \label{eq:JF4jt}
\end{equation}

The identity holds

\begin{definition}
Evolution of stable job employment 
\begin{equation*}
SJ_{jt} = SJ_{jt-1} + A_{4jt} - S_{4jt}
\end{equation*}
\end{definition}

\paragraph{Average stable  employment}

for establishment $j$ between periods $t-1$ and $t$%
\begin{equation}
\bar{SJ}_{jt}=\frac{\left( SJ_{jt-1}+SJ_{jt}\right) }{2}  \label{eq:sjbarjt}
\end{equation}

and appropriate job creation and destruction definitions can be derived.


\subsection{Caveats}

The above is defined in terms of end-of-quarter jobs. The published numbers
on the website are beginning-of-quarter jobs. Confusion may arise.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "lehd_definitions_revisions"
%%% End: 
